
---
title: "Do Conventional Strawberries Benefit Producers More Than Consumers?"
author: "Suakshi Soni"
date: "2025-05-09"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: cosmo
---

## Introduction

Strawberry farming in the United States involves significant chemical input, particularly in conventional systems. California and Florida—key producers—exhibit distinct chemical use and production outcomes. Using USDA NASS Census and Survey data, this analysis investigates whether producers benefit more than consumers from intensified chemical usage.

## Research Question

**Are producers the primary beneficiaries of intensified chemical use in strawberry farming, while consumers bear disproportionate health and financial costs?**

## Data & Methodology

### Data Selection and Filtering

```{r}
library(tidyverse)
library(ggplot2)

strawberry <- read_csv("strawb_mar6.csv", show_col_types = FALSE)

# Filter by Program and State
census <- strawberry |> filter(Program == "CENSUS")
survey <- strawberry |> filter(Program == "SURVEY")

cen_f <- census |> filter(State == "FLORIDA")
cen_c <- census |> filter(State == "CALIFORNIA")
sur_f <- survey |> filter(State == "FLORIDA")
sur_c <- survey |> filter(State == "CALIFORNIA")

# Chemical use only
strawb_sur_c <- sur_c |> filter(str_detect(`Domain`, "CHEMICAL"))
```

### Chemical Classification

```{r}
strawb_sur_c <- strawb_sur_c |> mutate(commas = str_count(`Data Item`, ","))

strawb_sur_c_3 <- strawb_sur_c |> filter(commas == 3)

# Split by commas
strawb_sur_c_3 <- strawb_sur_c_3 |> separate(`Data Item`, c("Fruit", "Category", "Item", "Metric"), sep = ",", extra = "merge")
strawb_sur_c_3 <- strawb_sur_c_3 |> separate(`Domain Category`, c("Domain_1", "Domain Category"), sep = ":", extra = "merge")

# Further split domain for chemical name
strawb_sur_c_3 <- strawb_sur_c_3 |> separate(Domain, c("Domain", "Chemical"), sep = ",", extra = "merge")

# Clean Value column
strawb_sur_c_3 <- strawb_sur_c_3 |> filter(Value != "(D)", Value != "(NA)")
strawb_sur_c_3$Value <- as.numeric(strawb_sur_c_3$Value)
```

## Chemical Usage Analysis

```{r}
# Summary by Year and Chemical Type
summary_by_year_chemical <- strawb_sur_c_3 |> group_by(Year, Chemical) |> summarise(Total_Value = sum(Value, na.rm = TRUE)) |> pivot_wider(names_from = Chemical, values_from = Total_Value)

# Boxplot
ggplot(strawb_sur_c_3, aes(x = as.factor(Year), y = Value, fill = Chemical)) +
  geom_boxplot() +
  labs(title = "Chemical Usage in California (2021–2023)", x = "Year", y = "lb/acre") +
  theme_minimal()
```

## Producer Income Analysis

```{r}
income_cen_c <- cen_c |> filter(Commodity == "INCOME, NET CASH FARM")
income_cen_f <- cen_f |> filter(Commodity == "INCOME, NET CASH FARM")

# Clean Value
income_cen_c$Value <- as.numeric(gsub(",", "", income_cen_c$Value)) / 1000
income_cen_f$Value <- as.numeric(gsub(",", "", income_cen_f$Value)) / 1000

# Density Plot for California
ggplot(income_cen_c, aes(x = Value)) +
  geom_density(fill = "lightblue", alpha = 0.5) +
  labs(title = "Net Income Distribution: California Producers", x = "Income ($ thousands)", y = "Density") +
  theme_minimal()

# Density Plot for Florida
ggplot(income_cen_f, aes(x = Value)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Net Income Distribution: Florida Producers", x = "Income ($ thousands)", y = "Density") +
  theme_minimal()
```

## Conclusion

This analysis finds that **California producers** gain significantly higher incomes from conventional strawberry farming compared to Florida. This is associated with **greater chemical input**, notably fungicides and insecticides.

Meanwhile, **consumers bear hidden costs**, including health risks from residue and high prices for organic alternatives. This underscores a need for **transparency, regulatory reform, and equitable food systems**.

## Limitations & Future Research

- Limited organic data made comparison difficult
- Consumer price and health exposure data were not included
- Future work can incorporate toxicity levels, consumer surveys, and explore the impact of regulation

## References

- USDA NASS (2023). Census and Survey Data.  
- Environmental Working Group (2023). "Dirty Dozen" Report.  
- CDC (2020). National Report on Chemical Exposure.  
- EPA (2021). Pesticide Use in Agriculture.  
- UC Davis (2014). Strawberry Cost Studies.
